---
title: "RStudio::conf18 - Tidy Eval"
author: "Julien Brun, NCEAS"
date: "ecodatascience, 2/28/2018"
output:
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
```

## Tidy eval

### Programming with dplyr, tidyr, and ggplot2

This talk was given by Hadley Wickham at the RStudio Conference 2018. 

I could not find the slides of his talk online, so this summary is based on the vignette on the same topic: http://dplyr.tidyverse.org/articles/programming.html


## What's different?

You might have noticed that when using the tidyverse, contrary to R base, you do **not** quote the column names. This is because most of the tidyverse functions use non-standard evaluation (NSE) and has its custom way to evaluate expressions.

. . . 

This has two main advantages:

- Operations on data frames can be expressed with a compact syntax
- This provide more flexibility in regards to when and how the evaluation occurs

## Some Drawbacks?

- You can't replace a value with a seemingly equivalent object that you've defined elsewhere (can't use variables directly)
- dplyr code can be ambiguous depending what is defined where

*=> this make programming with dplyr  not straight forward*

## Example Time

```{r example, echo=TRUE, message=FALSE}
library(dplyr)

# Define a dummy data frame
my_df <- as_tibble(iris)
my_df

```

## Computing the mean by species

```{r group_by_filtering, echo=TRUE}

my_df %>%
  group_by(Species) %>%
  summarise_all(mean, na.rm = TRUE)

```

## Using a function

```{r function time, echo=TRUE, eval = FALSE}
# Function definition
mean_by_group <- function(df, group_attr){
  my_df %>%
    group_by(group_attr) %>%
    summarise_all(mean, na.rm = TRUE)
}

# Function call
mean_by_group(my_df, Species)

# Maybe this?
mean_by_group(my_df, "Species")
```
## Our new tools

- *The quosure `quo()`*: to replace the quotation (to be exact it creates a special type of formula)
- *The bang `!!`*: to signal the need of unquoting

## Example

```{r function working, echo=TRUE}
# Function definition
mean_by_group <- function(df, group_attr){
  df %>%
    group_by(!!group_attr) %>%
    summarise_all(mean, na.rm = TRUE)
}

# The right call
mean_by_group(my_df, quo(Species))

```

## On cars?
```{r function working on cars, echo=TRUE}
# Can we use it on cars?
mean_by_group(mtcars, quo(cyl))
```
